---
title: "Dissolved Oxygen Observations"
subtitle: "across Long Island"
author: "Adam Starke"
date: "September 18, 2015"
output: html_document
theme: flatly
highlight: pygments

---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(shiny)
library(shinydashboard)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(lubridate)
library(shiny)
library(ggvis)
library(dygraphs)
library(xts)
library(magrittr)
library(dataRetrieval)
library(leaflet)
library(RColorBrewer)
library(rgdal)
library(sp)
library(maptools)
library(metricsgraphics)
library(htmltools)

# load('.RData')


```


```{r DataLoad, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}


## Start Pulling in USGS Site data	
# Programatic attempt- using a list of sites to add to all subsequent processes
# sites <- c('404327073072701', '404149073051301', '404153073030701')


siteIDs <- c('405004073391001', 
	     '405240073314901', 
	     '405232073281801', 
	     '405318073250101', 
	     '405329073212601', 
	     '403739073401001', 
	     '403731073353801', 
	     '403734073153401', 
	     '404513072555701', 
	     '404338073082101', 
	     '404327073072701', 
	     '404149073051301', 
	     '404153073030701')


siteNames <- c('Hemp', 
	       'OBH', 
	       'CSH', 
	       'Hunt', 
	       'NPH', 
	       'HB', 
	       'MB', 
	       'FI', 
	       'BP', 
	       'NCB', 
	       'GSB1', 
	       'GSB2', 
	       'GSB3')

names(siteIDs) <- siteNames


DOparameter <- "00300"


RawDO <- readNWISuv(siteIDs, DOparameter, "", "")

# Old Code ---- 
# <- '405004073391001'
# OBH <- '405240073314901'
# CSH <- '405232073281801'
# Hunt <- '405318073250101'
# NPH <- '405329073212601'
# HB <- '403739073401001'
# MB <- '403731073353801'
# FI <- '403734073153401'
# BP <- '404513072555701'
# NCB <- '404338073082101'
# GSB1 <- '404327073072701'
# GSB2 <- '404149073051301'
# GSB3 <- '404153073030701'
# 



# create a function that is given a site ID and outputs a 'raw data' file that is named according to the site the data is pulled from.



# # Site codes
# site1 <- sites[1]
# site2 <- sites[2]
# site3 <- sites[3]
# 
# northport <- '405329073212601'
# 
# # Parameter code



# 
# 
# site1_Rawdata <- readNWISuv(site1, DOparameter, "", "")
# site2_Rawdata <- readNWISuv(site2, DOparameter, "", "")
# site3_Rawdata <- readNWISuv(site3, DOparameter, "", "")
# northport_Rawdata <- readNWISuv(northport, DOparameter, "", "")
# 
# # Add spatial info and query time (to mark the time of the generation of the data)
# 
# site1_loc <- attr(site1_Rawdata, "siteInfo")
# site2_loc <- attr(site2_Rawdata, "siteInfo")
# site3_loc <- attr(site3_Rawdata, "siteInfo")
# northport_loc <- attr(northport_Rawdata, "siteInfo")

# Location data ----

siteInfo <- attr(RawDO, "siteInfo")

siteLocations <- siteInfo %>% 
	select(station_nm, site_no, starts_with("dec")) %>% 
	rename(Lat = dec_lat_va, Lon = dec_lon_va)

# 
# siteLocations <- bind_rows(site1_loc, site2_loc, site3_loc, northport_loc) %>% 
# 	select(station_nm, site_no, starts_with("dec")) %>% 
# 	rename(Lat = dec_lat_va, Lon = dec_lon_va)

bluepoints <- readOGR(".","BluepointsProperty", encoding = "ESRI Shapefile")

bluepoints <- spTransform(bluepoints, CRS("+init=epsg:4326"))

```
## Study Area
```{r Maps, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=10, fig.height=5}

DO_map <- leaflet(siteLocations) %>% 
	addProviderTiles("CartoDB.Positron") %>%
	addMarkers(popup = ~htmlEscape(station_nm), options = popupOptions(zoomAnimation = TRUE, closeOnClick = FALSE)) %>% 
	addPolygons(data = bluepoints, fill = FALSE, weight = 3) 
DO_map

```

## Data is preliminary
### Data has been corrected for Salinity and Fouling 
Data associated with this page can be found the following pages:  
http://waterdata.usgs.gov/nwis/inventory?agency_code=USGS&site_no=404149073051301  
http://waterdata.usgs.gov/nwis/inventory?agency_code=USGS&site_no=404327073072701  
http://waterdata.usgs.gov/nwis/inventory?agency_code=USGS&site_no=404153073030701  
http://waterdata.usgs.gov/nwis/uv?site_no=405329073212601

Each of the three stations we've established consist of 2 sets of Dissolved Oxygen loggers "Upper" loggers are set at 0.5 meters from the sediment and "Lower" loggers are set at 0.1 meters from the sediments.


```{r, echo=FALSE, fig.align='center', fig.width=9, fig.height=3}
# 
# # Function for dyplots
# 


DOdy_singlemonitor <- function(data){
	data %>%
	rename(DO_0.5m = X_00300_00011) %>%
	select(dateTime, starts_with("DO")) %>%
	xts(order.by = .$dateTime) %>%
	dygraph(., main = "Station 1- Mouth of Connetquot River", group = "DO-stations") %>%
	dyAxis("y", valueRange = c(-1, 11)) %>%
	dySeries('DO_0.5m', 'DO 0.5m') %>%
	dyRangeSelector(retainDateWindow = FALSE) %>%
	dyOptions(useDataTimezone = FALSE)
}


RawDO %>% filter(site_no == siteIDs[3:5]) %>%
	DOdy_singlemonitor(.)%>% dyShading(from = 2.3, to = -1, axis = "y", color = "#ffd1d1") %>%
	dyShading(from = 2.3, to = 4.8, axis = 'y', color = "#ffe8d1")


# 
# 
# DO1 <- site1_Rawdata %>% 
# 	rename(DO_0.5m = X_0.5m.above.seabed_00300_00011) %>% 
# 	rename(DO_0.1m = X_0.1m.above.seabed_00300_00011) %>% 
# 	select(dateTime, starts_with("DO")) %>% 
# 	xts(order.by = .$dateTime) %>% 
# 	dygraph(., main = "Station 1- Mouth of Connetquot River", group = "DO-stations") %>% 
# 	dyAxis("y", valueRange = c(-1, 11)) %>% 
# 	dySeries('DO_0.5m', 'DO 0.5m') %>% 
# 	dySeries('DO_0.1m', 'DO 0.1m') %>%
# 	dyRangeSelector(retainDateWindow = FALSE) %>% 
# 	dyOptions(useDataTimezone = FALSE, colors = RColorBrewer::brewer.pal(12, "Paired")[1:2])
# 
# 
# DO2 <- site2_Rawdata %>% 
# 	rename(DO_0.5m = X_0.5m.above.seabed_00300_00011) %>% 
# 	rename(DO_0.1m = X_0.1m.above.seabed_00300_00011) %>% 
# 	select(dateTime, starts_with("DO")) %>% 
# 	xts(order.by = .$dateTime) %>% 
# 	dygraph(., main = "Station 2- Central Bluepoints Property", group = "DO-stations") %>% 
# 	dyAxis("y", label = "DO mg/L", valueRange = c(-1, 11)) %>% 
# 	dySeries('DO_0.5m', 'DO 0.5m') %>% 
# 	dySeries('DO_0.1m', 'DO 0.1m') %>%
# 	dyRangeSelector(retainDateWindow = FALSE) %>% 
# 	dyOptions(useDataTimezone = FALSE, colors = RColorBrewer::brewer.pal(12, "Paired")[3:4])  
# 	
# 
# DO3 <- site3_Rawdata %>% 
# 	rename(DO_0.5m = X_0.5m.above.seabed_00300_00011) %>% 
# 	rename(DO_0.1m = X_0.1m.above.seabed_00300_00011) %>% 
# 	select(dateTime, starts_with("DO")) %>%
# 	xts(order.by = .$dateTime) %>% 
# 	dygraph(., main = "Station 3- Along Eastern Bluepoints Property Line", group = "DO-stations") %>% 
# 	dyAxis("y", valueRange = c(-1, 11)) %>% 
# 	dySeries('DO_0.5m', 'DO 0.5m') %>% 
# 	dySeries('DO_0.1m', 'DO 0.1m') %>%
# 	dyRangeSelector(retainDateWindow = FALSE) %>% 
# 	dyOptions(useDataTimezone = FALSE, colors = RColorBrewer::brewer.pal(12, "Paired")[9:10]) 
# 	
# 
# 
# DO1
# DO2
# DO3
# 

```

  
### Putting this in context with the EPA Oxygen Criteria:


Taken from: http://water.epa.gov/scitech/swguidance/standards/criteria/aqlife/dissolved/dofacts.cfm  

> What are the criteria limits?  
Dissolved oxygen criteria apply to both continuous and cyclic low DO conditions. If the DO conditions are always **_above_ the chronic criterion for growth (4.8 mg/L)**, the aquatic life at that location should not be harmed. If the DO conditions at a site are **_below_ the juvenile/adult survival criterion (2.3 mg/L)**, there is not enough DO to protect aquatic life. When persistent DO conditions are between these two values, further evaluation of duration and intensity of low DO is needed to determine whether the level of oxygen can support a healthy aquatic life community.  
The approach for episodic or cyclic low DO conditions requires that the DO be directly measured over a daily cycle or that the daily DO cycle be estimated. The cyclic pattern is then compared to allowable levels to determine suitability of DO conditions for the juvenile/adult survival, larval growth, and larval recruitment endpoints. 

Below are the same plots as above with these criteria higlighted. 

```{r, echo=FALSE, fig.align='center', fig.width=9, fig.height=3}
# 
# 
# DO1 %>% dyShading(from = 2.3, to = -1, axis = "y", color = "#ffd1d1") %>% 
# 	dyShading(from = 2.3, to = 4.8, axis = 'y', color = "#ffe8d1")
# 
# DO2 %>% dyShading(from = 2.3, to = -1, axis = "y", color = "#ffd1d1") %>% 
# 	dyShading(from = 2.3, to = 4.8, axis = 'y', color = "#ffe8d1")
# 
# DO3 %>% dyShading(from = 2.3, to = -1, axis = "y", color = "#ffd1d1") %>% 
# 	dyShading(from = 2.3, to = 4.8, axis = 'y', color = "#ffe8d1")



```

```{r, echo=FALSE, fig.align='center', fig.width=9, fig.height=3}

rankDOdata <- site2_Rawdata %>% 
	select(dateTime, X_0.5m.above.seabed_00300_00011, X_0.1m.above.seabed_00300_00011) %>% 
	mutate(CumDist_0.5m = cume_dist(X_0.5m.above.seabed_00300_00011), CumDist_0.1m = cume_dist(X_0.1m.above.seabed_00300_00011)) 

# 
# rankPlot <- rankDOdata%>% 
# 	ggvis(x = ~CumDist_0.5m, y = ~X_0.5m.above.seabed_00300_00011) %>% 
# 	layer_lines()
# rankPlot

CumDO <- rankDOdata %>% 
	mjs_plot(x = CumDist_0.5m, y = X_0.5m.above.seabed_00300_00011) %>% 
	mjs_line() %>%
	mjs_add_baseline(4.8, "Above this line, aquatic life unharmed") %>% 
	mjs_add_baseline(2.3, "Not Enough DO to protect life") %>% 
	mjs_axis_x(min_x = -0.10)
CumDO


```



